---
title: "Statistical Learning for Data Science Exam — Solutions"
output:
  html_document:
    toc: true
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  error = FALSE
)
```

The duration of the exam is 120 minutes. For most exercises you are
required to provide R code: write your code in the assigned R code
chunk. For some exercises you are required to provide a textual answer:
type your answer in the space that follows **Your answer**.

```{r prep}
matricola <- 
name <- 
surname <- 

if (!exists("matricola") || !exists("name") || !exists("surname") ||
    !is.numeric(matricola) || length(matricola) != 1 ||
    !nzchar(name) || !nzchar(surname)) {
  stop("Please insert valid student information in the first chunk.")
}
```

```{r packages, cache=FALSE}
library(ISLR)
library(tidyverse)
```

# Exercise 1

The `Wage` dataset in the ISLR package contains wage and demographic
information for 3000 male workers in the Mid-Atlantic region of the
United States. Each row represents one worker.

Consider the following variables:

* `year`: year in which the observation was recorded;
* `age`: worker's age;
* `education`: education level;
* `jobclass`: type of occupation, either industrial or information-based;
* `health`: worker's self-reported health status;
* `health_ins`: whether the worker has health insurance;
* `wage`: worker's wage index.

Load the data using:

```{r load-wage}
data(Wage)
```

## (a)

Determine:

* The number of observations and variables.
* The type of each variable listed above.
* Whether any of these variables contain missing values.

Briefly explain the difference between a numerical variable and a
categorical variable.

**Your answer:**

```{r ex1a}

```



## (b)

Create a new categorical variable named `age_group` with the following categories:

* `younger than 35` for workers younger than 35;
* `35–49` for workers aged between 35 and 49;
* `50 or older` for workers aged 50 or older.

Report the number of workers in each age group.

**Your answer:**

```{r ex1b}

```



## (c)

For each education level, calculate:

* the number of workers;
* the mean wage;
* the median wage;
* the standard deviation of wage.

Arrange the resulting table from the highest to the lowest mean wage.

Which education group has the highest average wage? Is the ordering of
the mean wages the same as the ordering of the median wages?

**Your answer:**

```{r ex1c}

```



## (d)

Construct a histogram of `wage`. Describe the shape of the distribution. Comment on its center,
variability, skewness, and any potentially unusual observations.

**Your answer:**

```{r ex1d}

 
```


## (e)

Construct side-by-side boxplots showing the distribution of `wage` for
each education level.

**Your answer:**

```{r ex1e}

```

## (f)

With respect to the previous point, comment on:

* differences in median wage;
* differences in wage variability;
* the presence of potential outliers.

Does the graph suggest that obtaining more education causes wages to
increase? Explain briefly.

**Your answer:**


## (g)

Construct a scatterplot of `wage` against `age`. Describe the
relationship.

**Your answer:**

```{r ex1g}

```



# Exercise 2

In this exercise, the objective is to construct and assess a model for
predicting wage.

## (a)

Fit a linear regression model with `wage` as the response variable and
`age` as the explanatory variable. Display the model summary, interpret
the coefficient of `age`, and comment on its statistical significance.

**Your answer:**

```{r ex2a}

```



## (b)

The relationship between `age` and `wage` might not be completely
linear. Fit a quadratic model and compare it with the linear model.
Which model do you prefer? Justify your answer.

**Your answer:**

```{r ex2b}

```



## (c)

Add `education` to the preferred model. State the reference category
used by R and interpret one of the coefficients associated with
education.

**Your answer:**

```{r ex2c}

```


## (d)

Add `jobclass` to the model from part (c). Compare the two models. Does
`jobclass` improve the model fit after accounting for age
and education?

**Your answer:**

```{r ex2d}

```



## (e)

Use the final model to predict the wage of a 40-year-old worker with an
education level of `4. College Grad` and job class `2. Information`.
Describe the difference between the point prediction and a prediction
interval.

**Your answer:**

```{r ex2e}

```



# Exercise 3

The `Smarket` dataset in the ISLR package contains daily percentage
returns for the S&P 500 stock market index between 2001 and 2005. Each
row represents one trading day.

Consider the following variables:

* `Year`: year in which the observation was recorded;
* `Lag1`: percentage return on the previous trading day;
* `Lag2`: percentage return two trading days earlier;
* `Volume`: daily trading volume;
* `Today`: percentage return on the current trading day;
* `Direction`: whether the market went `Up` or `Down` on the current
  trading day.

The objective is to predict market direction using information
available from previous trading days.

Load the data using:

```{r load-smarket}
data(Smarket)
```

## (a)

Determine:

* the number of observations and variables;
* the type of the response variable `Direction`;
* the number and proportion of observations for which the market went
  `Up` and `Down`.

Is the response variable approximately balanced?

**Your answer:**

```{r ex3a}

```



## (b)

Create a training set containing all observations from 2001 to 2004 and
a test set containing all observations from 2005.

Report the number of observations in each dataset.

Explain briefly why using earlier observations for training and later
observations for testing is appropriate in this application.

**Your answer:**

```{r ex3b}

```


## (c)

Using the training data, estimate a logistic regression model in which
`Direction` is the response variable and `Lag1` and `Lag2` are the input
variables.

Display the model summary.

Interpret the estimated coefficient of `Lag2`. Which input variables,
if any, are statistically significant at the 5% level?

**Your answer:**

```{r ex3c}

```


## (d)

Use the fitted logistic regression model to predict the probability that
the market goes `Up` for each observation in the test set.

Classify an observation as:

* `Up` if the predicted probability is greater than 0.5;
* `Down` otherwise.

Construct a confusion matrix comparing the observed and predicted
directions.

**Your answer:**

```{r ex3d}



```



## (e)

Calculate the classification accuracy of the logistic regression model
on the test set.

Next, consider a simple benchmark classifier that always predicts the
most frequent direction in the training set. Calculate the test accuracy
of this benchmark classifier.

Does the logistic regression model perform better than the benchmark?
Briefly comment on the predictive usefulness of the model.

**Your answer:**

```{r ex3e}

```



# Session Info

Do not modify the chunk below
```{r}
Sys.info()
```
